Type-Safe Modular Parsing

### 2017-06-10 SLE_17/sections/PackratParsing.tex
Revision

Moreover, the underlying parsing formalism should make backtracking
cheap, since \emph{backtracking is pervasive in modular parsing}.
Although we chose Packrat parsing (due to its immediate availability
and support), any other parsing formalism that provides similar
features should be adequate for modular parsing.

>>>>

Moreover, the underlying parsing formalism should make \emph{backtracking
cheap}, due to its pervasiveness in modular parsing.
Although we chose Packrat parsing, any other parsing formalism that provides similar
features should be ok.

### 2017-06-10 SLE_17/sections/RelatedWork.tex
Revision

Our work touches upon several topics including extensible parsing,
parser combinators and extensibility techniques. However, as far as we
know there's no work that discusses how to do statically type-safe and
separately compilable modular parsing.

@@

\paragraph{Syntactically Extensible Parsing}
Extensible parser generators~\cite{antlr1995,Grimm2006,Gouseti2014,Viera2012,Warth2016}
are a mainstream area of modular syntax and parsing. They allow users to write
modular grammars, where new
non-terminals and production rules can be introduced, some can even
override existing rules in the old grammar modules. For instance,
\textit{Rats!}~\cite{Grimm2006} constructs its own module system for
the collection of grammars, while NOA~\cite{Gouseti2014} uses Java
annotation to collect all information before producing an ANTLR~\cite{antlr1995} grammar
and the parsing code. Those parser
generators focus on the \textit{syntactic extensibility} of grammars:
they rely on whole compilation to generate a global parser, even if
there is only a slight modification in the grammar. Some of those
parser generators may statically check the correctness and unambiguity
of grammars. In contrast, because our approach is based on parser
combinators, there is no support for ambiguity checking.  However, as
far as we are aware, no extensible parser generators support separate
compilation or modular type-checking. It is worth mentioning that in~\cite{Viera2012}, users can define
grammar fragments as typed Haskell values, and combine them on the fly. Later they are processed
by a typed parser generator. Nevertheless, it only support separate compilation
on grammar fragments; a whole compilation is still needed for an extended grammar. And it requires
a lot of advanced language features, making client code less expressive. Our approach is simple from a
straightforward use of OO programming, and makes parsing code directly reusable.

>>>>

Our work touches upon several topics including extensible parsing,
parser combinators and extensibility techniques. 
%However, as far as we
%know there's no work that discusses how to do statically type-safe and
%separately compilable modular parsing.

@@

\paragraph{Syntactically Extensible Parsing}
Extensible parser generators~\cite{antlr1995,Grimm2006,Gouseti2014,Viera2012,Warth2016}
are a mainstream area of modular syntax and parsing. They allow users to write
modular grammars, where new
non-terminals and production rules can be introduced, some can even
override existing rules in the old grammar modules. For instance,
\textit{Rats!}~\cite{Grimm2006} constructs its own module system for
the collection of grammars, while NOA~\cite{Gouseti2014} uses Java
annotation to collect all information before producing an ANTLR~\cite{antlr1995} grammar
and the parsing code. Those parser
generators focus on the \textit{syntactic extensibility} of grammars:
they rely on whole compilation to generate a global parser, even if
there is only a slight modification in the grammar. Some of those
parser generators may statically check the correctness and unambiguity
of grammars. In contrast, because our approach is based on parser
combinators, there is no support for ambiguity checking.  However, as
far as we are aware, no extensible parser generators support separate
compilation or modular type-checking. It is worth mentioning that in~\cite{Viera2012}, users can define
grammar fragments as typed Haskell values, and combine them on the fly. Later they are processed
by a typed parser generator. Nevertheless this requires
a lot of advanced language features, making client complex. Our
approach is simple and a straightforward use of OO programming, 
and makes parsing code directly reusable.
